---
title: "The title"
author:
- name: First Author
  affiliation: '1'
  corresponding: yes
  address: Postal address
  email: my@email.com
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Ernst-August Doelle
  affiliation: '1,2'
  role: Writing - Review & Editing
shorttitle: Title
output:
  html_document:
    df_print: paged
authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
keywords: keywords
wordcount: X
bibliography: r-references.bib
toc: yes
toc_float: yes
floatsintext: no
figurelist: no
tablelist: no
footnotelist: no
linenumbers: no
mask: no
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Wilhelm-Wundt-University
- id: '2'
  institution: Konstanz Business School
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

<!--
Citations in papaja (delete appropriately later)

Add the bibtex entry in the .bib file. You can find the entries in Google scholar, 
but double check since it is not always correct. 

Call the citations in the text:

Citation within parentheses	[@R-papaja]

Multiple citations	[@R-papaja; @R-base]

In-text citations	@R-papaja

Year only	[-@R-base]

Only if your citation appears in the text it will also show up in the Reference 
list. Don't manually modify the Reference list.
-->

## Executive Summary
<!--
Executive Summary Description (150 words) – 0.3 POINTS

Summarize the report. Write this as the very last thing.

What is the main topic you are addressing?

What are your research questions and hypotheses?  

What are your results and the main conclusion?
-->

## Introduction 
<!--
Introduction Description (about 1000 words) – 0.5 POINTS

Place your topic of choice within the existing literature and explain what you 
are going to address in this report and why.

*	What is the main topic that is going to be studied in this paper?

*	Why is it important?

*	What are the existing studies that address it already?

*	What are your questions and hypotheses? Why are they important? How are they 
connected to previous work or to a problem you raised?

*	Which methods are you using? Why?

*	How is your work going to contribute to the field/or to whom (e.g., industry)?

*	What does the rest of the report look like? In one short paragraph, list the 
topics of the following sections.
-->
#### topic of choice and litterature research

The data used for this project comes from an open source provided by the European Union (EU). Lots of studies are conducted such as [reference above]. The mentioned studies are examples of how Twitter statuses of members of the European Parliament (MEP) can be connected to predict voting. Coalitions and political party cohesion are strong indicators of which vote will be final for the given MEP. The European Parliament is constructed in such a way where one political coalition may not have the overall majority to vote in or out a resolution. It is made on purpose to encourage discussion, debate and legal amendments. What tips the voting in favour or against are the members that vote against what is “expected” from them, in the sense of against of their priors. Having an idea of when or how these events happen are interesting to investigate from where this paper. The current literature has a similar methodology and investigative scope than our project. Our paper will corroborate or not with their findings. Towards the end it will be explained why it is more nuanced.
Our work is going to contribute to the field by providing additional insight into the voting behaviors of given MEPs. Our data being recent (2021) can suggest whether voting patterns are staying similar in time in comparison to other studies such as [reference from list]. 

#### Methods
The programming language used for this project is R. After having done some exploratory network descriptive to see whether there are some outliers or unusual dynamics going around. The following characteristics were studied:
•	Network density
•	Average path length 
•	Diameter
•	Degree census
•	Global transitivity
The Exponential family random graph models (ERGM) are our main tool for this project.
There are numerous studies that used this method such as:
• Gondal, The local and global structure of knowledge production in an emergent research field, SOCNET 2011 
• Lomi and Palotti, Relational collaboration among spatial multipoint competitors , SOCNET 2011 
• Wimmer & Lewis, Beyond and Below Racial Homophily, AJS 2010 
• Lusher, Masculinity, educational achievement and social status, GENDER & EDUCATION 2011 
• Rank et al. (2010). Structural logic of intra-organizational networks, ORG SCI, 2010.
ERGMs are Increasingly being used, increasingly being understood, increasingly being able to handle imperfect data (also missing link prediction) - Lusher, Koskinen, Robins ERGMs for SN, CUP, 2011
Quadratic assignment procedure (QAP) is another method that will be used for this project. It scrambles the dependent variable data through several permutations. By taking the data, and “scrambling” it repeatedly, resulting in multiple random datasets with the dependent variable—and then multiple analyses can be performed. It is like a bootstrapping method but applied in the field of network analysis. 

2-Does political cohesion, country of origin can predict the vote of a MEP against or for a given resolution ?
Method:
Linear regression based on metric extracted from network could be interesting. 

Hypotheses:
	Cohesion is stable across the two bills.
	Networks have little change from one bill to the other.
	MEPs do not agree with each other across bills.
	Political parties have little influence on the community formation.
	Countries have little influence on the community formation identified using transitivity.
	Are voting networks built around cycles grouping certain indviduals.
	What happens to the network when the most central nodes are taken away. This hypothesis would test whether the removal of a central node would be influential or not.


#### Roll-call votes
The analysis conducted in this paper is carried out over a set of 61 roll-call votes from two motions for European Parliament resolutions concerning different political areas of interest (45 from a climate-related motion and 16 from an industry-related motion). Roll-call votes are votes conducted in the European parliament where the vote (or absence of vote in some cases) of an MEP is recorded along with their name. Roll-call votes are only held under certain conditions, for example, when requested by a significant number of MEPs or if it is impossible to determine a majority from a show of hands \citealt{cherepnalkoski_karpf_mozetič_grčar_2016}. These votes are documented in the minutes of that day’s parliament session and are accessible on the European parliament site.

#### Impact of climate change on vulnerable population

As the COP26 is upon us between the 1st and 12th of November 2021 in Glasgow, climate change is becoming more and more central to international political agendas. New goals will be set to protect those most vulnerable to climate change. In fact, climate change leading to an increase of sea levels may turn out to be deadly for some marginalized societies such as the Sentinelese. A couple centimeters of additional water level can lead their islands to be submerged and permanently lose their foot on the solid ground. Other societies are under the increasing threat of higher temperatures leading to a whole spectrum of climate and weather changes. This may force certain communities to change the crop that is being grown within the region. Bad harvests may lead some societies to starvation or economical crisis due to bad quality products thus unfit for exportation towards richer areas such as Europe. Climate change affect many countries in many ways. Ecosystems are being disturbed create new geopolitical tensions such as access to drinkable water, new mineral and fossil fuel accesses in the arctic, sea commercial areas wanting to be extended for commercial reasons. All this without even mentioning the consequence on wildlife with their variety decreasing each year. As the president of Brazil would claim, what is the economical gain for a country to have all this wildlife in comparison to burning it for farmland. In conclusion, the communities that are already on the edge of survival will be once again hit hard by the result of a planet-degrading, unsustainable consumer behavior lived in the richest countries of the World. To push this trend further, emerging countries such as China are more and more willing to adopt such a way of living creating even more stress on the planet's carbon emission. 

#### Proceeding of report's parts
Firstly the methodology is explained. Some general comments are made on the obtention of the data, exploratory results under the form of plots. 
Secondly, an explanation of the exact manner of how the network and its modelling is constructed.
Thirdly, an interpretation of the modelling results of the networks, how it relates and partially answers the research question given previously.

## Methodology

### Dataset 
<!-- 
Dataset Description (about 500 words) 1 POINT (+ BONUS)

*	Which data set are you going to use? Three options:

  *	Use readily/easily available data (0 bonus points)

  *	Combine two or more existing datasets (max 0.5 bonus points)

  *	Scrape or collect your own data (max. 1 bonus point)

*	Clearly explain where the data is coming from:

  *	Who collected the data?

  *	What is the source?

  *	When was the data produced?

  *	How was the data collected?

*	Provide descriptive measures of your data (tables, plots, etc.)


*	Why is this data useful to study your topic and answer your research questions?

*	What is the potential bias in the data? How does this affect your results?
-->
As the dataset for this research two votes of the European Parliament on May 18th 2021 are scraped from the VoteWatch.eu website. The first vote is named "The impacts of climate change on vulnerable populations in developing countries", and has 44 subvotes. The second vote is named "Research Fund for Coal and Steel", and has [TODO] subvotes. The full official document of the votes can be found in the Appendix [TODO]. VoteWatch is a non-profit organization with the goal of providing insight into the European Parliament and their decision making for European citizens. VoteWatch describes their methodology on their website to great extend \cite{votewatch_methodology}. The foundation, however, lays in them automatically scraping the the voting data from the European Parliaments own website.


#### Source of data and collection
The data used in this research was collected by scraping the VoteWatch website for all votes concerning the two relevant motions. VoteWatch is a European political analysis website that aims to make the workings of the European parliament more transparent \citealt{votewatch_europe}. They therefore replicate the data from European parliament votes in a more easily accessible manner. It was therefore possible via this interface to scrape the data using a Python script with the urllib library. The name of the MEP, their European-level political party, member state and action relating to that particular vote (for, against, didn’t vote, abstain, absent) were collected via this method and collated to give two datasets with all actions concerning the two motions.

#### Usefulness of data for research questions and potential biases
The collected data concerns the names of MEPs, their nationality, European-level political party and vote for each relevant motion for resolution. This data provides enough information to be able to answer the research questions discussed above as the effect of political party (and nationality?) on agreement between individuals can be modelled. The collection of data from two motions in different fields also provides the opportunity to be able to compare the network of agreement between the two areas and comment on similarities and differences. The creation of two networks from these two datasets is discussed in the next section of this paper.
A limitation of the data collected is the lack of information about the personal network dynamics at play in the European parliament. Agreement between individuals is likely also influenced by who they are friends with and collaborate with most often. There are likely also more powerful members of the party such as national and European-party level leaders who may therefore invoke a higher level of agreement from others. With the data collected these dynamics are either impossible to model or can only be done in a limited manner.
The data collected also allows the comparison of networks for a motion relating to the climate-area and a more industry focused motion. However, for this research only two motions were chosen rather than examining all relevant motions in the current period. This has been carried out in previous research, see \cite{cherepnalkoski_karpf_mozetič_grčar_2016} but was deemed to be too time consuming for this initial research. The research carried out in this paper is therefore considered to be a point-in-time comparison which could trigger future investigation if this brief analysis indicates interesting results.
An issue with the collection method of the roll-call data is that it was collected via a third-party website VoteWatch rather than the European parliament website. This decision was taken due to the significant difference in ease in navigating and accessing the data on the two sites. The risk of this choice is considered to be small as the aim of VoteWatch is transparency so the data is likely to be accurate but it is not impossible that inaccuracies and differences could have been caused when reproducing the vote results between the two websites. Given more time to conduct this research, a next step would be to build a method for finding the relevant votes in the European parliament minutes and scraping the data directly.







### Data analysis (Research Rationale)
<!--
Data Analysis (Research Rationale) (about 500 words) – 1 POINTS

*	Why are these two methods suitable for your data?

*	Why are these two methods suitable for your research questions?

*	Are there other methods to address these questions? If yes, why are the 
methods you chose better for this case?
-->

The first research question looks into the correlation between the network from the climate bill and the network from the industry bill. We propose to use a QAP correlation test to find an answer for this question. QAP is similar to CUG, in that it uses statistical simulation to generate a distribution of hypothetical networks. However, with QAP we do not generate a series of random graphs (with some properties such as a given size and density), but we now control for all purely structural properties of the two graphs being compared themselves. This is done by making permutations of the networks that are to be compared, similar to the practice of bootstrapping.

A QAP correlation test is the most suitable option. A QAP test is suitable for situations where two or more networks have to be compared with each other. For the first research question, two networks, based on votes from two different bills, have to be compared. A CUG model would not work here, since it is only about investigating the network measures of one network. Similarly, an LNAM model focuses on the social influence in a network and is not applicable. It might be possible to make two ERGM models for the bills, where explanatory variables are included that represent groups that vote alike, but this would be needlessly complex and inferior to the QAP test. Lastly, a MRQAP model is not necessary since only two networks have to be compared. 

The second research question investigates whether there is cohesion of agreement, based on the MEP’s voting behaviour for the climate bill, within every European group. In the climate bill network two MEP’s are connected if they agree on an arbitrary amount of amendments within the bill. Since we want to explain the reason why a tie is formed in the network or specifically between two MEP’s, we propose to use an exponential random graph family model or ERGM for short. An ERGM is a class of network models to perform causal inference and deduce the cause of an observed relation phenomenon that we observe and conceptualize as a network. The model will include endogenous and exogenous terms since the party membership is an attribute of the vertices and therefore a term like nodefactor needs to be used. Nodefactor is a Markovian term that measures the extent to which being in a certain category influences edge formation. We specifically need to use the regular variant as opposed to the directed in or directed out variant since the climate bill network is undirected. The complexity and flexibility of ERGM makes this technique very suitable to verifying this research question. The other techniques are not suitable for this type of analysis since CUG and QAP are used for checking whether a network statistic is statistically significant and comparing two networks respectively.

## Results
<!--
Results Description (about 2000 words)
-->

#### Descriptive measures

In order to get a better overview of the dataset, the five-number summary is computed for both the climate motion and the industry motion as a function of the agreement threshold. It was chosen to develop the plots as a function of the agreement threshold because the dataset for both motions is very large. There are 705 members of the European parliament with 248160 undirected edges for just the climate motion. As a result, the edge density for the climate motion is equal to 1 for the complete dataset. Each edge between two parliament members if formed when they agree on a minimum of 1 vote within the motion. Therefore not all edges are a result of causal relationships. In order to visualise just the important, non-random relationships, this section will explore the descriptive measures as a result of the agreement threshold that is yet to be determined. 

```{r}
par(mfrow = c(2,2), cex=0.6)
mep_climate <- read.csv("../../data/mepdata_climate.csv", header = T, sep = ",")

edges_climate <- read.csv("../../data/edges_for_against.csv", header = T, sep = ",")

df <- data.frame(agreement=integer(), density=double())
for (i in 1:44) {
  edges_filter <- edges_climate[edges_climate[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_climate, directed=FALSE)
  density_filter <- igraph::transitivity(network_filter, type = 'global')
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Global transitivity", main="Transitivity as a function of agreement threshold")


df <- data.frame(agreement=integer(), density=double())
for (i in 1:44) {
  edges_filter <- edges_climate[edges_climate[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_climate, directed=FALSE)
  density_filter <- igraph::edge_density(network_filter)
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Network density", main="Network density as a function of agreement threshold")

df <- data.frame(agreement=integer(), density=double())
for (i in 1:44) {
  edges_filter <- edges_climate[edges_climate[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_climate, directed=FALSE)
  density_filter <- igraph::mean_distance(network_filter)
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Average path length", main="Average path length a function of agreement threshold")

df <- data.frame(agreement=integer(), density=double())
for (i in 1:44) {
  edges_filter <- edges_climate[edges_climate[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_climate, directed=FALSE)
  density_filter <- igraph::diameter(network_filter)
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Diameter", main="Diameter as a function of agreement threshold")


par(mfrow = c(2,2), cex=0.6)
mep_coalsteel <- read.csv("../../data/coal_steel_data.csv", header = T, sep = ",")

edges_coalsteel <- read.csv("../../data/edges_for_against_coal_steel.csv", header = T, sep = ",")


df <- data.frame(agreement=integer(), density=double())
for (i in 1:15) {
  edges_filter <- edges_coalsteel[edges_coalsteel[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_coalsteel, directed=FALSE)
  density_filter <- igraph::transitivity(network_filter, type = 'global')
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Global transitivity", main="Transitivity as a function of agreement threshold")


df <- data.frame(agreement=integer(), density=double())
for (i in 1:15) {
  edges_filter <- edges_coalsteel[edges_coalsteel[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_coalsteel, directed=FALSE)
  density_filter <- igraph::edge_density(network_filter)
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Network density", main="Network density as a function of agreement threshold")

df <- data.frame(agreement=integer(), density=double())
for (i in 1:15) {
  edges_filter <- edges_coalsteel[edges_coalsteel[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_coalsteel, directed=FALSE)
  density_filter <- igraph::mean_distance(network_filter)
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Average path length", main="Average path length a function of agreement threshold")

df <- data.frame(agreement=integer(), density=double())
for (i in 1:15) {
  edges_filter <- edges_coalsteel[edges_coalsteel[, "Agreement"]>i, c("MEP_1","MEP_2")]
  network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_coalsteel, directed=FALSE)
  density_filter <- igraph::diameter(network_filter)
  de <- list(agreement=i, density=density_filter)
  df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Diameter", main="Diameter as a function of agreement threshold")


```

##### Climate Motion

The graphs for the climate motion show a rather dense network with a short average path length. This can be explained by the fact that with the full dataset, almost all members of the EP are connected to each other through some vote agreement. Which seems logical, given that they would just need to agree on one vote. However, this is not the most relevant information to us. Instead, it is observed from the transitivity measure figure that at an agreement threshold of 15, there is a drop in transitivity. This signals that there is a reduction in clusters when a threshold of 15 is set. Similarly, the network density plot shows a vast decline around the threshold of 15, indicating a large reduction of edges. As such we assume that by setting the treshold slightly past 15 will eliminate random edges between members of the EP. It is thus an interesting threshold to further evaluate. 


##### Industry Motion




#### Creation of the networks
To create a network it first had to be determined what counted as agreement between two MEPs. Multiple actions can be taken as a response to a roll-call vote. While it is clear that two MEPs are in agreement if they both vote for or both against a motion, this is less clear in the case of abstaining, being absent or just not voting. This could be due to disagreement with the motion, perhaps an unwillingness to go against your party but could also be due to a personal issue, logistical problem or lack of information. It is therefore not possible to say if these MEPs have the same opinion about the vote. An agreement is therefore only considered valid in this research if they both voted for or against. This decision is in line with the edges created in previous research \citealt{cherepnalkoski_karpf_mozetič_grčar_2016}.
Furthermore, due to limitations with the ERGM’s ability to model weighted edges, the network to be used in this analysis had to consist of binary edges. Therefore, in order to create a network of agreement between MEPs a decision was required to determine what conditions would create an edge between a pair of MEPs. Previous research has created networks from each individual roll-call vote where an edge equals agreement in that particular vote \citealt{cherepnalkoski_karpf_mozetič_grčar_2016}. However, being able to make conclusions from this method requires an ERGM model to be created for each roll-call vote and then the results combined using a meta-analysis method. Therefore, given the time restrictions of this project, a decision was made to create an edge between individuals based on agreement in a certain number of votes or more following the approach adopted in previous papers, see \cite{cranmer_desmarais_2011}. 
An appropriate threshold therefore had to be determined. If the threshold would be too low then it would be more likely to cause degenerate graphs as most MEPs are likely to agree with each other on at least a few occasions leading to very dense graphs. On the other hand, if a too high threshold is selected then interesting network dynamics could be omitted by only including edges between MEPs who almost completely agree with each other. A balance therefore had to be found by generating the networks for a selection of thresholds, comparing density and the resulting ERGM model.
An exploration of the thresholds was started by plotting a histogram of the different levels of agreements between MEPs and then a line graph showing the changing density of the networks depending on the threshold for agreement required for an edge (see plots below). The two plots corroborate each other, with big changes in network density occurring when setting the threshold above approximately 15 and again above 40. This is caused by the two peaks that we can see present in the spread of agreement around these two points.

```{r}
edges_climate <- read.csv("../../data/edges_for_against.csv")
mep_climate <- read.csv("../../data/mepdata_climate.csv")
hist(edges_climate$Agreement, col = "lightblue", xlab = "Agreement",main='Histogram of spread of agreement between MEPs')
df <- data.frame(agreement=integer(), density=double())
for (i in 1:44) {
        edges_filter <- edges_climate[edges_climate[, "Agreement"]>i, c("MEP_1","MEP_2")]
        network_filter <- igraph::graph_from_data_frame(edges_filter, vertices=mep_climate, directed=FALSE)
        density_filter <- igraph::edge_density(network_filter)
        de <- list(agreement=i, density=density_filter)
        df = rbind(df,de, stringsAsFactors=FALSE) 
}

plot(df, type="l", col="blue", xlab="Agreement Threshold", ylab="Density", main="Changing density of graphs depending on agreement threshold")
```
For clarity the thresholds considered here are absolute numbers of votes on which the MEPs agree on. Since every MEP has the opportunity to be both present and vote on each motion for each resolution then this is considered to be a clear and fair method for testing levels of agreement.
With this data in mind, three models were created, visualised and tested with the ERGM (see plots and degree distribution of selected networks below). The three models used a threshold of above 5, above 20 and above 40 respectively. These models were all separately tested to ensure the best chance of finding a converging model. It is also possible that different dynamics might be in play in the model for occasional agreement (above 5) in comparison to almost total agreement (above 40) which can be tested with these three models. 

```{r}
#Filter to relevant edges
edges_climate5 <- edges_climate[edges_climate[, "Agreement"]>5, c("MEP_1","MEP_2")]
edges_climate20 <- edges_climate[edges_climate[, "Agreement"]>20, c("MEP_1","MEP_2")]
edges_climate40 <- edges_climate[edges_climate[, "Agreement"]>40, c("MEP_1","MEP_2")]
#Create network
network_climate5 <- igraph::graph_from_data_frame(edges_climate5, vertices=mep_climate, directed=FALSE)
network_climate20 <- igraph::graph_from_data_frame(edges_climate20, vertices=mep_climate, directed=FALSE)
network_climate40 <- igraph::graph_from_data_frame(edges_climate40, vertices=mep_climate, directed=FALSE)
#Plot network
par(mfrow=c(1,2))
plot(network_climate5,
     vertex.size = 3,
     edge.arrow.size = .4,
     vertex.label=NA,
     layout = igraph::layout.fruchterman.reingold,
     main= "Threshold above 5" )
hist(igraph::degree(network_climate5), xlab="Degree", main='Degree distribution above 5')
par(mfrow=c(1,2))
plot(network_climate20,
     vertex.size = 3,
     edge.arrow.size = .4,
     vertex.label=NA,
     layout = igraph::layout.fruchterman.reingold,
     main= "Theshold above 20")
hist(igraph::degree(network_climate20), ylim=c(0,600), xlab="Degree", main='Degree distribution above 20')
par(mfrow=c(1,2))
plot(network_climate40,
     vertex.size = 3,
     edge.arrow.size = .4,
     vertex.label=NA,
     layout = igraph::layout_with_graphopt,
     main= "Threshold above 40")
hist(igraph::degree(network_climate40), ylim=c(0,600), xlab="Degree", main='Degree distribution above 40')
```

### Model 1
<!--
Model 1 Description (about 1000 words) – 2.5 POINTS

*	Present your results appropriately (plots, tables…) and discuss your findings 
in plain English

*	Discuss the meaning of your findings in relation to your hypothesis. (half of
the points evaluated in this other part)
-->

```{r echo = FALSE, results = 'asis'}
# table example
cn <- c("age", "gender", "eyes_col")
one <- c(7, "M", "BLUE")
two <- c(8, "F", "BROWN")
three <- c(8, "M", "GREEN")
four <- c(7, "F", "PINK")

tab <- rbind(cn, one, two, three, four)
rownames(tab) <- NULL
knitr::kable(tab)
```

### ERGM 
<!--
ERGM Description (about 1000) – 2.5 POINTS

* Present your results appropriately (plots, tables…) and discuss your findings 
in plain English

* Discuss the meaning of your findings in relation to your hypothesis. (half of 
the points evaluated in this other part)
-->

Option 1: 

```{r model 1, echo = FALSE}
# model results display example
ctl <- c(4.17,5.58,5.18,6.11,4.50,4.61,5.17,4.53,5.33,5.14)
trt <- c(4.81,4.17,4.41,3.59,5.87,3.83,6.03,4.89,4.32,4.69)
group <- gl(2, 10, 20, labels = c("Ctl","Trt"))
weight <- c(ctl, trt)
lm.D9 <- lm(weight ~ group)
lm.D90 <- lm(weight ~ group - 1)

# install.packages("texreg")

knitr::kable(texreg::matrixreg(lm.D9))


```

Option 2

```{r model 2, echo = FALSE, results = 'asis'}
# model results display example
ctl <- c(4.17,5.58,5.18,6.11,4.50,4.61,5.17,4.53,5.33,5.14)
trt <- c(4.81,4.17,4.41,3.59,5.87,3.83,6.03,4.89,4.32,4.69)
group <- gl(2, 10, 20, labels = c("Ctl","Trt"))
weight <- c(ctl, trt)
lm.D9 <- lm(weight ~ group)
lm.D90 <- lm(weight ~ group - 1)

# install.packages("texreg")

knitr::kable(texreg::matrixreg(list(lm.D9, lm.D90)))


```


Option 3

```{r model 3, echo = FALSE}
# model results display example
ctl <- c(4.17,5.58,5.18,6.11,4.50,4.61,5.17,4.53,5.33,5.14)
trt <- c(4.81,4.17,4.41,3.59,5.87,3.83,6.03,4.89,4.32,4.69)
group <- gl(2, 10, 20, labels = c("Ctl","Trt"))
weight <- c(ctl, trt)
lm.D9 <- lm(weight ~ group)
lm.D90 <- lm(weight ~ group - 1)

# install.packages("texreg")


texreg::plotreg(lm.D9)


```


Option 4

```{r model 4, echo = FALSE}
# model results display example
ctl <- c(4.17,5.58,5.18,6.11,4.50,4.61,5.17,4.53,5.33,5.14)
trt <- c(4.81,4.17,4.41,3.59,5.87,3.83,6.03,4.89,4.32,4.69)
group <- gl(2, 10, 20, labels = c("Ctl","Trt"))
weight <- c(ctl, trt)
lm.D9 <- lm(weight ~ group)
lm.D90 <- lm(weight ~ group - 1)

# install.packages("texreg")

texreg::plotreg(list(lm.D9, lm.D90))

```


## Conclusion
<!--
Conclusion Description (about 350 words) – 0.7 POINTS

What were your topic and research questions again? (1 sentence)

What did you learn from the two analysis you run? *** most important point to 
address 0.5 POINTS here

Who benefits from your findings?

What does remain an open problem?

Can you give suggestions for future work in this area?
-->

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
